Article data analysis

### 2019-01-15 notebooks/article_notebook.Rmd
updated with mlst data and adjusted a few figures

## Further work

- Assemble genomes

- Align assemblies

- Land on methodology for phylogenetic analysis (likely core genome SNP - ML)

@@

@@

mut_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, gene_names) %>%
    group_by(key, value, species) %>%
    count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(1)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(1)
    ) +
    scale_fill_manual(values = palette) +
    labs(y = "Percent (%) of isolates") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 90,
        size = 12,
        hjust = 1,
        vjust = 0.4
      ),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~ species)
```

@@

```{r, fig.width=14, fig.height=16,message = FALSE, warning = FALSE}
# # identify which sequence types are represented over three times
# n_ST <- mlst_data[c("id","ST")] %>%
#   group_by(ST) %>%
#   count() %>%
#   filter(n > 3)
# 
# # species ST  
# species_st <- mlst_data[c("id","ST")] %>%
#   left_join(isolate_data[c("id", "species")], by = "id") %>%
#   select(-id) %>%
#   group_by(ST, species) %>%
#   count() %>%
#   filter(ST %in% n_ST$ST)
# 
# # prepare micplot data for each ST with n > 3
# micplot_st_df <- mlst_data[c("id", "ST")] %>%
#   left_join(mic_data[c("id", "CIP")], by = "id") %>%
#   group_by(ST, CIP) %>%
#   count() %>%
#   filter(ST %in% n_ST$ST)
# 
# # prepare heatmap data frame
# st_heatmap_df <- mlst_data[c("id", "ST")] %>%
#   left_join(mut_report[c("id", "marR")], by = "id") %>%
#   left_join(cols_to_add, by = "id") %>%
#   left_join(acquired_report, by = "id") %>%
#   select(-id) %>%
#   mutate_at(vars(-ST),
#             funs(as.numeric)) %>%
#   group_by(ST) %>%
#   summarise_all(funs(sum, n = n())) %>%
#   ungroup() %>%
#   rename("n" = marR_n) %>%
#   select(-contains("_n")) %>%
#   gather(mechanism, value, -c(ST, n)) %>%
#   mutate(mechanism = sub("_sum", "", mechanism),
#          percent = round(value/n*100),
#          value2 = ifelse(value > 0, 1, 0)) %>%
#   filter(ST %in% n_ST$ST)
# 
# # plots
# factor_breaks <- c(0.015,0.03,0.06,0.12,0.25,0.5,1,2,4,8,16,32,64,128,256)
# factor_levels <- c("0.015","0.03","0.06","0.12","0.25","0.5","1","2","4","8","16","32","64","128","256")
# 
# ## heatmap
# heatmap_st <- ggplot(st_heatmap_df, aes(mechanism, factor(ST), fill = percent))+
#   geom_tile()+
#   scale_fill_viridis(option = "A")+
#   guides(fill = FALSE)+
#   theme(axis.text.x = element_text(angle = 90,
#                                    hjust = 1,
#                                    vjust = 0.3,
#                                    size = 8),
#         axis.title.x = element_blank())
# 
# ## micplot
# micplot_st <- ggplot(micplot_st_df, aes(factor(ST), CIP))+
#   geom_boxplot(fill = "#67a9cf")+
#   scale_y_continuous(labels = factor_levels,
#                      breaks = factor_breaks,
#                      trans = "log10")+
#   theme(axis.title = element_blank(),
#         axis.text.y = element_blank(),
#         axis.text.x = element_text(size = 8),
#         axis.ticks.y = element_blank())+
#   coord_flip()
# 
# # n plot
# nplot_st <- ggplot(species_st, aes(factor(ST), n, fill = species))+
#   geom_col(color = "black")+
#   scale_fill_manual(values = palette) +
#   theme(axis.title = element_blank(),
#         axis.text.y = element_blank(),
#         axis.text.x = element_text(size = 8),
#         axis.ticks.y = element_blank(),
#         legend.position = c(0.6, 0.95),
#         legend.title = element_blank())+
#   coord_flip()
# 
# cowplot::plot_grid(heatmap_st, micplot_st, nplot_st, ncol = 3, align = "h", rel_widths = c(3/6, 2/6, 1/6))

@@

gheatmap(rotated_fan_tree,
         total_gene_data,
         offset = 0.005,
         width = 0.2,
         colnames_offset_y = 3,
         colnames_position = "top",
         font.size = 2.5) +
  scale_fill_manual(values = palette3,
                    labels = c("Absent",
                               "Acquired gene present",
                               "Intrinsic gene mutated"))
```

>>>>

@@

## Sequence types
```{r}
mlst_data[,c("id","ST")] %>%
  group_by(ST) %>%
  count()
```

## Sequence types per animal species
```{r}
mlst_data[,c("id","ST")] %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(ST, species) %>%
  count()
```

@@

mut_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, gene_names) %>%
    group_by(key, value, species) %>%
    count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    mutate(species = factor(species,
                            levels = c("Broiler","Pig","Wild bird","Red fox"))) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(1)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(1)
    ) +
    scale_fill_manual(values = palette) +
    scale_x_discrete(limits = c("gyrA","gyrB","parC","parE","marA","marR","acrB","rpoB")) +
    labs(y = "Percent (%) of isolates") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 90,
        size = 12,
        hjust = 1,
        vjust = 0.4
      ),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~ species)
```

@@

@@

final_tree <- gheatmap(rotated_fan_tree,
         total_gene_data,
         offset = 0.005,
         width = 0.2,
         colnames_offset_y = 3,
         colnames_position = "top",
         font.size = 2.5) +
  scale_fill_manual(values = palette3,
                    labels = c("Gene/Mutation absent",
                               "Acquired gene present",
                               "Intrinsic gene mutated"))
final_tree
```
